Arrows for Parallel Computations

### 2017-10-25 paper/content/benchmarks.tex
minor

We have tested our parallel DSL and algorithmic skeletons implemented
in it. Benchmarks were conducted both in a shared and in a distributed
memory setting. All benchmarks were done on the ``Glasgow grid'', consisting of
16 machines with 2 Intel\SymbReg~Xeon\SymbReg~E5-2640 v2 and 64 GB of DDR3 RAM each. Each processor has 8 cores and 16 (hyper-threaded) threads with a base frequency of 2 GHz and a turbo frequency of 2.50 GHz. This results in a total of 256 cores and 512 threads for the whole cluster. The operating system was Ubuntu 14.04 LTS with Kernel 3.19.0-33. Non-surprisingly, we found that hyper-threaded 32 cores do not behave in the same manner as real 16 cores (numbers here for a single machine). We disregarded the hyper-threading ability in most of the
cases.

@@

The difference between, say, PArrows with |Par| Monad backend and a
genuine |Par|
Monad benchmark is very small. To give an example, it is $0.4\seconds$ in favour of PArrows for 16 cores ($10.8\seconds$ vs. $11.2\seconds$) and $-0.8\seconds$ in favour of the |Par| monad for 8 cores ($16.1\seconds$ vs. $16.9\seconds$) for
the Sudoku benchmark in the shared memory setting. It is almost invisible in speedup and
(non shown) run time plots. We thus show only the results for the
PArrows-enabled versions.

To show that PArrows induce very small overhead in a distributed context as well, we compare the original Eden
versions of the benchmark to its PArrows-enabled counterpart in the \rmtest, \torustest and \jacobitest benchmarks. We plot execution time differences between measurements for
PArrows and the corresponding backend in a separate plot
(Figs.~\ref{fig:bench-rm-dist}--\ref{fig:torusBenchmark}). As an example, the differences range in
about $0.5\seconds$ for the execution time of $46\seconds$ on 256 cores
for distributed \rmtest with PArrows and Eden. For these comparisons, the plots show absolute
time differences that are not relative \wrt the total execution time.
Furthermore, the error bars ends were computed from point-wise maximum of both standard
deviations from both measurements for PArrows and non-PArrows
versions. These are the values provided by the |bench| package that we
used for bench-marking. We call a difference between two versions
significant when the border of the error bar of absolute time
difference is above or below zero. In other words: the time
difference is significant if it is above measurement error.

@@

In the shared memory setting we naturally find that the backends perform differently. Furthermore, in our distributed memory tests on the full grid the difference between PArrows and Eden was almost always below the error margin if PArrows performed worse. Even the biggest difference of 12.27\% in the \jacobitest for input $2^{4253}-1$ and 128 cores remained in line with these findings. The only exception to this was the \torustest, which showed only a $1.7\%$ difference in favour of Eden which is only a marginal slowdown in our opinion.

>>>>

In the following we describe the benchmarks of our parallel DSL and algorithmic skeletons we conducted. They were run both in a shared and in a distributed
memory setting. All benchmarks were done on the \enquote{Glasgow grid}, consisting of
16 machines with 2 Intel\SymbReg~Xeon\SymbReg~E5-2640 v2 and 64 GB of DDR3 RAM each. Each processor has 8 cores and 16 (hyper-threaded) threads with a base frequency of 2 GHz and a turbo frequency of 2.50 GHz. This results in a total of 256 cores and 512 threads for the whole cluster. The operating system was Ubuntu 14.04 LTS with Kernel 3.19.0-33. Non-surprisingly, we found that hyper-threaded 32 cores do not behave in the same manner as real 16 cores (numbers here for a single machine). We disregarded the hyper-threading ability in most of the
cases.

@@

The difference between, say, PArrows with |Par| Monad backend and a
genuine |Par|
Monad benchmark is very small. To give an example, it is $0.4\seconds$ in favour of PArrows for 16 cores ($10.8\seconds$ vs. $11.2\seconds$) and $-0.8\seconds$ in favour of the |Par| monad for 8 cores ($16.1\seconds$ vs. $16.9\seconds$) for
the Sudoku benchmark in the shared memory setting. It is almost invisible in speedup and
(non shown) run time plots. We thus show only the results for the
PArrows-enabled versions in the backend-comparison.

To show that PArrows induce very small overhead even in a distributed context (and in general as well), we compare the original Eden
versions of the benchmark to its PArrows-enabled counterpart in the \rmtest, \torustest and \jacobitest benchmarks. We plot execution time differences between measurements for
PArrows and the corresponding backend in a separate plot
(Figs.~\ref{fig:bench-rm-dist}--\ref{fig:torusBenchmark}). As an example, the differences range in
about $0.5\seconds$ for the execution time of $46\seconds$ on 256 cores
for distributed \rmtest with PArrows and Eden. For these comparisons, the plots show absolute
time differences that are not relative \wrt the total execution time.
Furthermore, the error bars ends were computed from point-wise maximum of both standard
deviations from both measurements for PArrows and non-PArrows
versions. These are the values provided by the |bench| package that we
used for bench-marking. We call a difference between two versions
significant when the border of the error bar of absolute time
difference is above or below zero. In other words: the time
difference is significant if it is outside of the measurement error.

@@

In the shared memory setting we naturally find that the backends perform differently. Furthermore, in our distributed memory tests on the full grid the difference between PArrows and Eden was almost always below the error margin if PArrows performed worse. Even the biggest difference of 12.27\% in the \jacobitest for input $2^{4253}-1$ and 128 cores remained in line with these findings. The only exception to this was the \torustest  with only a $1.7\%$ difference in favour of Eden which is only a marginal slowdown in our opinion.
